In-Band Synchronization for\\Distributed SDN Control Planes\thanks{Supported by the European Institute of Innovation \& Technology (EIT) project 13153 (Software Defined Networking) and the FP7 EU project UNIFY.}

### 2015-11-17 ccr-camera/inband-sync.tex
some comments and fixes

The meta-configuration is implemented as a set of flow entries
designed in a way that it does not affect the processing of data plane
traffic. 
We will later discuss later how this can be achieved, for example in 
today's OpenFlow.

@@

%\petr{How to make sure that traffic is not affected by the ``meta'' part?}
\vspace{1mm}
\noindent\textbf{Synchronization.}
 %
We provide the controllers with the ability to execute
sequences of \emph{operations} on the
switch configuration in an \emph{atomic} (transactional) or \emph{non-atomic} mode.
In the atomic mode, either \emph{all} operations in the sequence, henceforth simply
called \emph{transaction},
should be executed, or none (the transaction is rejected).
In the non-atomic mode, no atomicity guarantees are provided.

@@

To read the current configuration of a switch (the existing flow entries), the controller should send
the OpenFlow \emph{add flow monitor} command with the \textsf{ofpt\_initial} optional flag.
In this paper, we write \textbf{read-state} for sending
this message, receiving a response, and returning the received configuration.
\petr{Is it atomic? Should we say that the command returns a ``snapshot'' of the
  configuration?}

@@

We now describe Algorithms~\ref{alg:claim}-\ref{alg:compare} in more
detail.
Each of the operations generates one or two FlowMod commands.
The \textit{match} part in these commands specifies only the $64$bit meta-data
field in the header space, leaving the remaining fields at default
values.
In defining the value and the mask,
% of the math part,
we
use the notation $a\concat b$ to represent a concatenation of two
 integers $a$ and $b$: $a\concat b := (a<<s)+b$, where $s$ is the size of $b$ in bits.
We also write \texttt{$\ones{s}$} (\texttt{$\zeros{s}$}) for  a $s$ bit 
string of $1$'s ($0$'s).
%
The action part in these commands is defined as an integer, which in
practice can be implemented by a \emph{set meta-data} instruction where the written value is that integer.

@@

\vspace{1mm}
\noindent\textbf{Write and compare.}
The memory \memwrite and \compare
 operations are implemented using similar techniques.
To write a value $k$ in a memory address $\addr$ (Algorithm~\ref{alg:write}), a new flow entry is
added, in which the value part of the match field contains $\addr$ and
the mask part contains $k$. 
Therefore, the effective match string od the entry 
 is $k'\concat \addr$, where $k'$ is result os replacing every zero bit in $k$ with a wildcard ($*$).
  In order to rewrite a memory location, the old entry needs to be deleted and a new one need to be added atomically.
  %Values written to the same address will replace one another,
%following the command definition to replace the old flow entry with a new one in case they
%share the same match parts.
Similarly to claim entries, we set the action to a magic value,
$\memmagic$, in order to distinguish them from policy entries.
%[[PK not needed?
%returned by \textbf{read-state}.
%]]

@@

@@

A less trivial Algorithm~\ref{alg:update} adresses the problem of
policy identifiers growing without bound in  
Algorithm~\ref{alg:simple-update} by, additionally, using our id-claimer abstraction.
%to bound the number of used policy identifiers.
Here, the controller first reads the current policy id and 
claims it, which prevents another controller from using the id 
for a different policy (Lines~\ref{update:read-state1}-\ref{update:read-state2}).Then the controller claims the first \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}).
Finally, within an atomic transaction (\textbf{\execatomic}), it verifies whether the chosen policy id has not
been claimed and that
the policy has not been concurrently modified, and then updates the
policy.
%[[PK 
%We observe that our  algorithm is only
%\emph{lock-free}, as it allows scenarios in which only one live controller
%succeeds in installing its updates. We can easily make it
%wait-free by reading the state of the switch and claiming the current
%policy id within the same {\execatomic} construct.
%]]

@@

We observe that our implementations of synchronization primitives
(Algorithms~\ref{alg:claim}-\ref{alg:compare}) 
%is \emph{switch-space efficient}
%in the sense that 
require only a constant number of flow entries per memory address.
The number of used addresses is constant for  
Algorithm~\ref{alg:simple-update} and linear in the number of
controllers for Algorithm~\ref{alg:update}. 
%[[PK not constant: proportional to the number of failures. But this
%is kind of obvious.
%Also, in case of controller failures, at most a constant amount of space is lost.
%]]

@@

To investigate the feasibility of our approach,
we implemented a proof-of-concept of our algorithms
as a python library which wraps ovs-ofctl, 
a well-known command line tool for monitoring and administering OpenFlow switches. 
The code is publicly available at \url{https://github.com/lironsc/of-sync-lib}.

>>>>

The meta-configuration is implemented as a set of flow entries
designed in a way that it does not affect the processing of data plane
traffic. 
We will later discuss later how this can be achieved, for example in 
today's OpenFlow. \liron{two times "later"!}

@@

%\petr{How to make sure that traffic is not affected by the ``meta'' part?}
\vspace{1mm}
\noindent\textbf{Synchronization.}
 %
We provide the controllers with the ability to execute
sequences of \emph{operations} on the
switch configuration in an \emph{atomic} (transactional) or \emph{non-atomic} mode.
In the atomic mode, either \emph{all} operations in the sequence, henceforth simply
called \emph{transaction},
should be executed, or none (the transaction is rejected).
In the non-atomic mode, no atomicity guarantees are provided.
\liron{all these jumps from atomic to transactional and back is confusing...}

@@

To read the current configuration of a switch (the existing flow entries), the controller should send
the OpenFlow \emph{add flow monitor} command with the \textsf{ofpt\_initial} optional flag.
In this paper, we write \textbf{read-state} for sending
this message, receiving a response, and returning the received configuration.
\petr{Is it atomic? Should we say that the command returns a ``snapshot'' of the
  configuration?}\liron{I am not sure if this is atomic, considering that we execute other things in a bundle then this is practically atomic for us. Perhaps we can leave it unspecified as it is now?}

@@

We now describe Algorithms~\ref{alg:claim}-\ref{alg:compare} in more
detail.
Each of the operations generates one or two FlowMod commands.
The \textit{match} part in these commands specifies only the $64$bit meta-data
field in the header space, leaving the remaining fields at default
values.\liron{this is not accurate, the match part simply considers only the metadata field. no need to discuss other fields, in any case matching doesn't change fields and doesn't assume any values to other fields.}
In defining the value and the mask,
% of the math part,
we
use the notation $a\concat b$ to represent a concatenation of two
 integers $a$ and $b$: $a\concat b := (a<<s)+b$, where $s$ is the size of $b$ in bits.
We also write \texttt{$\ones{s}$} (\texttt{$\zeros{s}$}) for  a $s$ bit 
string of $1$'s ($0$'s).
%
The action part in these commands is defined as an integer, which in
practice can be implemented by a \emph{set meta-data} instruction where the written value is that integer.

@@

\vspace{1mm}
\noindent\textbf{Write and compare.}
The memory \memwrite and \compare
 operations are implemented using similar techniques.
To write a value $k$ in a memory address $\addr$ (Algorithm~\ref{alg:write}), a new flow entry is
added, in which the value part of the match field contains $\addr$ and
the mask part contains $k$. 
Therefore, the effective match string of the entry 
 is $k'\concat \addr$, where $k'$ is result of replacing every zero bit in $k$ with a wildcard ($*$).
  In order to rewrite a memory location, the old entry needs to be deleted and a new one need to be added atomically.
  %Values written to the same address will replace one another,
%following the command definition to replace the old flow entry with a new one in case they
%share the same match parts.
Similarly to claim entries, we set the action to a magic value,
$\memmagic$, in order to distinguish them from policy entries.
%[[PK not needed?
%returned by \textbf{read-state}.
%]]

@@

\liron{SECOND TRY: Note that comparing an address for the first time, without writing it first, always succeeds and also writes to it the compared value. In case where such comparison may occur to some address $addr$, it is advised to first perform a safe initialization by calling $\textit{\compare}(\addr,0)$ which fails in case that $addr$ has already been used (and the initialization is not desired).}

@@

A less trivial Algorithm~\ref{alg:update} adresses the problem of
policy identifiers growing without bound in  
Algorithm~\ref{alg:simple-update} by, additionally, using our id-claimer abstraction.
%to bound the number of used policy identifiers.
Here, the controller first reads the current policy id and 
claims it, which prevents another controller from using the id 
for a different policy (Lines~\ref{update:read-state1}-\ref{update:read-state2}). Then the controller chooses an \emph{unused} id and computes the update commands to be executed (Line~\ref{update:compute}). \liron{I have replaced "claims the first" with "chooses an" following Petr email}
Finally, within an atomic transaction (\textbf{\execatomic}), it verifies whether the chosen policy id has not
been claimed and that
the policy has not been concurrently modified, and then updates the
policy.
%[[PK 
%We observe that our  algorithm is only
%\emph{lock-free}, as it allows scenarios in which only one live controller
%succeeds in installing its updates. We can easily make it
%wait-free by reading the state of the switch and claiming the current
%policy id within the same {\execatomic} construct.
%]]

@@

We observe that our implementations of synchronization primitives
(Algorithms~\ref{alg:claim}-\ref{alg:compare}) 
%is \emph{switch-space efficient}
%in the sense that 
require only a constant number of flow entries per memory address or claim.
The number of used addresses is constant for  
Algorithm~\ref{alg:simple-update} and linear in the number of
controllers for Algorithm~\ref{alg:update}. 
%[[PK not constant: proportional to the number of failures. But this
%is kind of obvious.
%Also, in case of controller failures, at most a constant amount of space is lost.
%]]

@@

To investigate the feasibility of our approach,
we implemented a proof-of-concept of our algorithms
as a python library which wraps ovs-ofctl, 
a well-known command line tool for monitoring and administering OpenFlow switches. 
The code is publicly available at \url{https://github.com/lironsc/of-sync-lib}. \liron{perhaps we can put it as reference? (putting the link in bib)}
